---
title: 'MLE: Lab 1'
author: "Andy Ballard"
date: "January 13, 2017"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{multirow}
- \usepackage{dcolumn}
---

## Data Structures

R has 5 main data structures that can be defined by a) the number of dimensions (1d, 2d, nd) and whether all items must be of the same type (homogenous) or can be of different types (heterogeneous).

$$\begin{tabular}{rrr}
  \hline
 & \emph{Homogenous} & \emph{Heterogeneous}\\ 
 \hline
 \emph{1d} & Atomic Vector & List \\
 \emph{2d} & Matrix & Data Frame \\
 \emph{nd} & Array & \\
  \hline
\end{tabular}$$


## Setting up your workspace

```{r, message=FALSE}
# Start with a clean workspace
rm(list=ls())

#Doing any simulations, randomization, sampling, etc? Set your seed
set.seed(1988) 

# Load libraries
library(foreign)
library(xtable)
library(apsrtable)
library(arm)
library(ggplot2)

```

What if we need to install packages and load them? The native R way is pretty simple...

```{r eval=FALSE}
install.packages("package.name.here")
library(package.name.here)
```

...but messy. You need to do this separately for each package. Instead, we can write our own function to 1) check if packages we specify are installed, 2) if not, install them, and 3) load the desired packages. Particularly when we want to load dozens of packages for complicated projects, this is a much easier way to do things.

```{r message=FALSE}
loadPkg=function(toLoad){
  for(lib in toLoad){
    if(! lib %in% installed.packages()[,1])
    { install.packages(lib, repos='http://cran.rstudio.com/') }
    suppressMessages( library(lib, character.only=TRUE) )
  }
}

toLoad <- c("foreign", "xtable", "apsrtable", "arm", "ggplot2")
loadPkg(toLoad)
```




```{r message=FALSE}

# Other functions that I use frequently
char = function(x){ as.character(x) }
num = function(x){ as.numeric(char(x)) }
detach_package <- function(pkg, character.only = FALSE)
{
  if(!character.only)
  {
    pkg <- deparse(substitute(pkg))
  }
  search_item <- paste("package", pkg, sep = ":")
  while(search_item %in% search())
  {
    detach(search_item, unload = TRUE, character.only = TRUE)
  }
}

```



# Fun with R

Monty Hall Monte Carlo

```{r}
doors <- c("1", "2", "3")

correct <- 0
incorrect <- 0

while((correct + incorrect) <= 10000){

prize <- sample(doors, 1)

guess <- sample(doors, 1)


if(guess != prize){
  one.removed <- c(guess, prize)
} else {
  one.removed <- c(guess, sample(doors[doors != guess], 1))
}

final.guess <- one.removed[one.removed != guess]

if(final.guess==prize){
  correct <- correct + 1
} else if (final.guess!=prize){
  incorrect <- incorrect + 1
}

}

correct/10000
incorrect/10000

```


Slides from today can be found [here](http://www2.stat.duke.edu/~cr173/Sta523_Fa14/slides/2014-09-03-data_struct.html#1). You can also find a wonderful post on R data structures [here](http://adv-r.had.co.nz/Data-structures.html). Happy hunting.
